Demography and linked selection in wild and domesticated maize

### 2015-10-06 manuscript.tex
updates

Maize represents an excellent organism to study these phenomena. Maize is a species of tremendous importance worldwide as both a staple crop \cite{shiferaw2011} and as a model for understanding plant evolution \cite{strable2009}. 
Broadly speaking, archaeological and genetic studies have
established that maize domestication is likely to have taken place in
 Mexico approximately 9,000 years bp
\cite{smith1995,matsuoka2002}. Teosinte, the most
recent wild ancestor to maize, remains extant throughout much of the
Americas \cite{wilkes1967}. Additionally, several large-effect
domestication loci \cite{doebley1995, wills2013, wang2015} and putative domestication
regions \cite{hufford2012} have been identified. But despite all that is
known about maize domestication, the parameters of the
domestication process remain uncertain. Specifically, the size of the
maize domestication bottleneck has not been estimated independently of
the bottleneck's duration, nor are there sequence-based estimates of the effective
population size of modern maize. Sequence information from maize and
teosinte may therefore be utilized to address these questions.

To that end, the objectives of our study were to 1) investigate the
relative importance of different forms of selection on whole-genome
variability, using maize and teosinte as a model 2) research the impact that the
demography history has had on genetic variability in maize, and how
this compares to the impact of a different demographic history in
teosinte; and 3) precisely estimate the parameters of the
maize domestication bottleneck.  \cite{chia2012}. \mbh{A quick sentence here about why this matters broadly would help...what kind of basic inferences will this research allow us to make about the interplay of demography and selection?}

\dropcap{D}omesticated plant species evolve in a unique fashion
compared to their wild counterparts \cite{doebley2006}.
This is a result of both the anthropomorphic nature of artificial selection on
domesticates \cite{purugganan2009} as well as the demographic characteristics of the domestication
bottleneck(s) that they tend to have experienced
\cite{ross2007}. However, the
complex interplay between selective pressures and demographic
limitations, and the impact that this interplay has on identifying
selection and understanding demography, is not fully understood. Although a large body of
research that involves searching the genomes of domesticated species for evidence
of positive selection exists \cite{hufford2012, he2011, vigouroux2002, chapman2008}, these studies tend to focus on
identifying or mapping particular genes or regions that play an
important role in phenotypic evolution. In contrast, knowledge regarding the impacts that demography and
selection have on whole-genome patterns of genetic variability remains limited.

>>>>

Maize represents an excellent organism to study these phenomena. Maize is a species of tremendous importance worldwide as both a staple crop \cite{shiferaw2011} and as a model for understanding plant evolution \cite{strable2009}. 
Archaeological and genetic studies have
established that maize domestication is likely to have taken place in
 Mexico approximately 9,000 years bp
\cite{smith1995,matsuoka2002}. Teosinte, the most
recent wild ancestor to maize, remains extant throughout much of the
Americas \cite{wilkes1967}. Additionally, several large-effect
domestication loci \cite{doebley1995, wills2013, wang2015} and putative domestication
regions \cite{hufford2012} have been identified. But despite all that is
known about maize domestication, the parameters of the
domestication process remain uncertain. Specifically, the size of the
maize domestication bottleneck has not been estimated independently of
the bottleneck's duration, nor are there sequence-based estimates of the effective
population size of modern maize. At the same time, the predominant form or forms of selection that have taken place during and since domestication remain unknown. Sequence information from maize and teosinte may therefore be utilized to address these questions.

To that end, the objectives of our study were to 1) investigate the
relative importance of different forms of selection on whole-genome
variability, using maize and teosinte as a model 2) research the impact that the
demography history has had on genetic variability in maize, and how
this compares to the impact of a different demographic history in
its wild progenitor teosinte; and 3) precisely estimate the parameters of the
maize domestication bottleneck. Gaining an improved understanding of the interaction between demography and selection in maize will help to increase the understanding of these phenomenon across a wide array of organisms.

%% \dropcap{D}omesticated plant species evolve in a unique fashion
%% compared to their wild counterparts \cite{doebley2006}.
%% This is a result of both the anthropomorphic nature of artificial selection on
%% domesticates \cite{purugganan2009} as well as the demographic characteristics of the domestication
%% bottleneck(s) that they tend to have experiencedtu
%% \cite{ross2007}. However, the
%% complex interplay between selective pressures and demographic
%% limitations, and the impact that this interplay has on identifying
%% selection and understanding demography, is not fully understood. Although a large body of
%% research that involves searching the genomes of domesticated species for evidence
%% of positive selection exists \cite{hufford2012, he2011, vigouroux2002, chapman2008}, these studies tend to focus on
%% identifying or mapping particular genes or regions that play an
%% important role in phenotypic evolution. In contrast, knowledge regarding the impacts that demography and
%% selection have on whole-genome patterns of genetic variability remains limited.
